Performance of two different ontological and data access methods

### 2014-07-02 cold2014/magnus.tex
mention alignment
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@709 43ef1190-90fd-4dd7-a769-d0121f02b80e
remove todo
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@710 43ef1190-90fd-4dd7-a769-d0121f02b80e
add discussion
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@711 43ef1190-90fd-4dd7-a769-d0121f02b80e

@@

% TODO differences between Daimler and Renault
% TODO Alignment

@@

>>>>

To align the above ontologies, we used SPARQL CONSTRUCT statements
from the lexicon to each specification.

@@

@@

\section{Discussion}

We will discuss the results in terms of \textit{representation} and
\textit{structure}. The structure is the ontology definitions from
classes to properties.  The representation is how the actual data is
represented with each ontology. For instance, how the fuel type
\textit{Diesel} is represented.

\subsection{The small scale experiments}
The first experiment, using only three factors, was done to simulate a
small workload on the ontologies. This can represent an application
that present the user for a specification one at the time. The
configurator Renault has on their web page does exactly
that.\footnote{\url{http://www.semanlink.net/2012/cold/configurator.html}}
Our application made the workload on both ontologies to be as equal as
possible.  The only difference from a programming perspective was that
with CO we had to move to a new configuration URI after each
specification choice.

In the Figure \vref{3factorGraph} we can see that only one significant
effects, the O factor. In the results data found in appendix
\ref{appD} we can see that the Configuration Ontology was almost two
seconds slower than the VSO/COO runs.  Three out of four CO runs kept
the response time around two seconds, but one run had to use
approximately 3.5 seconds which resulted in an average run time of
$\approx2.3$ seconds.  In a study done to research the amount of time
a user would be willing to wait, they found out that the tolerable
waiting time for any information retrieval was approximately 2
seconds.~\cite{waitTime} This means that with three factors we were
still within the time frame for a good user experience.

One reason that CO was slower than VSO/COO for this experiment is that
all the car models represented with CO is placed in a separate RDF
graph. Each model had their own lexicon with all the information.
This means that an application has to retrieve in all 22 RDF graphs,
the amount of cars Renault has present in their API. These 22
retrievals are done right after a HTTP post is detected. This differs
from how the cars in VSO/COO are represented. There it is one RDF
graph per base model, which means that there were only two RDF graphs
to retrieve.  In our application this results in 20 more RDF graph
retrievals with CO than VSO/COO and because the three factor
experiment was so small, it had a higher impact than in the later
experiments.

Another factor that would influence the results is the amount of
values the search has to iterate over.  In Table \vref{averages}, we
presented all the average values and the average number of values
present in the Daimler and Renault data. We can see that CO have
$0.54$ more fuel types on average to iterate over and $0.3$ more
transmissions. This is not a lot, but can account for some delay in
the response time.

\subsection{The large scale experiments}
The last two experiments we define as large scale experiments. This is
because the workload and time consumption were significantly larger
than the two previous experiments. Both of these experiments were done
to simulate a workload for an application were one could define all
the specifications at once. Defining all specifications means that a
user can query for all specifications that may be present in a product
model.

The last experiment done was the one with ten factors. This experiment
had four more factors than the six factor experiment.  We chose to add
factors which represented specifications that were present in every
car model in each RDF graph. That was because some specifications like
height was not always declared in a car model and then would skew the
results. These incidents were present in CO's RDF graphs.

In Figure \vref{10factorGraph} we can see the results from this
experiment. There are unfortunately a lot of illegible significant
factors, but these are the less significant ones. In Table
\vref{10factorEffect} we see the top twenty absolute effects and
similar to all the other experiments, O is on top of the list. The
average response time of a run done against CO was $\approx58.5$
seconds while a run against VSO/COO had an average response time of
$\approx0.3$ seconds.  This corresponds with the absolute effect of
the O factor to be 58 seconds.

If we take a look at the other factors, Fuel Consumption (FC) has
taken the top single factor spot on the list.  This can be explained
with the same reason as when emission was topping the list in the six
factor experiment. Each car has a lot of possible FC values to iterate
over. In CO, each car has an average amount of values of approximately
5.8. For VSO/COO it is 3.75.  This is the second highest overall after
emission. This by itself does not explain why FC is on the top of the
list, but when we take a look at when FC is queried it becomes more
clear. The order each specification is queried is shown below.  We can
see that FC comes right before E which means that it will have a
greater impact on the amount of values checked later.

 \begin{enumerate}

		\item Gears (G)
		\item Transmission (T)
		\item Fuel Consumption (FC)
		\item Emission (E)
		\item Doors (D)
		\item Total Weight (WT)
		\item Fuel Type (FT)
		\item Weight (W)
		\item Seating Capacity (SC)
\end{enumerate}

If we take another look at Table \ref{10factorEffect} we see that
there are only three single factors except O that are in the top 15
significant factors.  After the 15 significant factor the absolute
effect drops significantly. The three factors are E, FC and T. E and T
are present because of the same reasons as in the six factor
experiment. We also see in the top 15 that there are only these three
factors present in all the factor interactions as well, which
indicates that T, FC and E affects each other. This is logical because
if a car has a high fuel consumption it also has a high emission
rate. There are some possible explanations for why there are so many
significant factor interactions in some of the experiments. The order
will affect this as mention above. Another explanation is that we
chose an average value to be an upper and lower limit of a factor. We
used these as factor levels to avoid an unrealistic high number of
levels. This could have opened for more new specifications to check on
one level of the factor.  As we see in the Table
\vref{10factorEffect}, \textit{E:WT} is higher than \textit{WT}
itself. This might indicate that one of the levels of E will open up
for more values to traverse in one of the levels of WT. This is a
complex matter which will depend on each product model as well as the
interaction between the specifications.

\section{Pros and cons}
There are several ways of making an application handle PDPs. As
mentioned there can be differences in collecting data, presenting data
and using data. This means that a recommended ontology and data
representation needs to be robust. This means that it has to be able
to perform well under different scenarios.  The ideal semantic
situation would not rely on string comparison, but with both
ontologies one have to deal with string literals eventually. A string
literal is still the value a user is presented with.  It depends on
how much of the data one wants machine accessible and how much is for
internal human reading.

The ontologies in this thesis are not made for tackling every possible
problem, but can perform well within their domain. The ontology
structure of CO in conjunction of the data representation, is made for
one particular scenario. That is to specify one specification after
another, not all at once. Here CO has several tools at disposal to
make it easier trimming down the result sets. The property
\textit{co:impossible} was to no use in our application due that our
application specified all specifications before searching. This
property lets the application remove possibilities if a user have
chosen some specifications. One would still has to use the lexicon as
reference, but could eliminate further possibilities. The property
\textit{co:impliedSpec} does something similar and gives the user a
list of specifications that are implied and do not need to be searched
for.  COO have some of the same functions and possibilities, but some
of these has to be reasoned over to be useful. One can produce the
same functionality by using \textit{coo:incompatibleWith} with
reasoning.

From the performance tests and programming, the product specific
ontology came out as the best.  This, as discussed, is not solely due
to the ontology structure. In the generic ontology, many of the issues
lies within the representation.  The generic ontology can easily be
used to represent different PDPs which will let the user stick to one
ontology.  Some alterations to the representation is needed to make it
more robust than what Renault have done.  They mention that reasoning
is something a third party user should be shielded from. We disagree
because if the user can choose when the reasoning is done they can
avoid possible bottlenecks. This means that for the representation to
work properly we recommend using a similar compatibility system to
COO. Here one can add the triples to each RDF graph of the product
model and do reasoning here. This gives the user freedom to choose
whatever way of reasoning and storing they might find most applicable,
which could be to pre-compute any big reasoning operations that may be
unsuited for real time. This will remove all the extra queries which
has to be done to move through Renault's large graph of
configurations. This representation puts more workload on the third
party user, but it will most likely decrease the response time
drastically.
